---
title: "Gender Pay Gap"
---

```{r load-libraries, echo=FALSE}
library(tidyverse) 
library(mosaic)
library(ggthemes)
library(GGally)
library(readxl)
library(here)
library(skimr)
library(janitor)
library(broom)
library(tidyquant)
library(infer)
library(openintro)
library(tidyquant)
```
# Omega Group plc- Pay Discrimination


At the last board meeting of Omega Group Plc., the headquarters of a large multinational company, the issue was raised that women were being discriminated in the company, in the sense that the salaries were not the same for male and female executives. A quick analysis of a sample of 50 employees (of which 24 men and 26 women) revealed that the average salary for men was about 8,700 higher than for women. This seemed like a considerable difference, so it was decided that a further analysis of the company salaries was warranted. 

You are asked to carry out the analysis. The objective is to find out whether there is indeed a significant difference between the salaries of men and women, and whether the difference is due to discrimination or whether it is based on another, possibly valid, determining factor. 

## Loading the data


```{r load_omega_data}
omega <- read_csv(here::here("data", "omega.csv"))
glimpse(omega) # examine the data frame
```

## Relationship Salary - Gender ?

The data frame `omega`  contains the salaries for the sample of 50 executives in the company. Can you conclude that there is a significant difference between the salaries of the male and female executives?

Note that you can perform different types of analyses, and check whether they all lead to the same conclusion 

.	Confidence intervals
.	Hypothesis testing
.	Correlation analysis
.	Regression

Calculate summary statistics on salary by gender. Also, create and print a dataframe where, for each gender, you show the mean, SD, sample size, the t-critical, the SE, the margin of error, and the low/high endpoints of a 95% confidence interval

```{r, confint_single_valiables}

# Confidence intervals
# Hypothesis testing
# Correlation analysis
# Regression

# Summary Statistics of salary by gender
mosaic::favstats (salary ~ gender, data=omega)

# Dataframe with two rows (male-female) and having as columns gender, mean, SD, sample size,
# the t-critical value, the standard error, the margin of error, 
# and the low/high endpoints of a 95% confidence interval

gender_salary <- omega %>%
  group_by(gender) %>%
  summarise(Mean = mean(salary), 
            SD = sd(salary), 
            Count = n(), 
            T_critical = qt(p = 0.975, df = Count-1), 
            SE = SD/sqrt(Count),
            Margin = T_critical * SE,
            Lower = Mean - Margin,
            Upper = Mean + Margin)

gender_salary

```

> What can you conclude from your analysis? A couple of sentences would be enough

> **Given that the confidence interval's of both genders do not overlap, it means that the difference in executive female and male salaries is statistically significant.**

You can also run a hypothesis testing, assuming as a null hypothesis that the mean difference in salaries is zero, or that, on average, men and women make the same amount of money. You should tun your hypothesis testing using `t.test()` and with the simulation method from the `infer` package.

*H-null: difference in salaries between male and female = 0. H-alt: difference in salaries between male and female != 0.* 

```{r, hypothesis_testing on salary}
# hypothesis testing using t.test() 

t_gender <- t.test(salary ~ gender, data = omega)
t_gender
t_gender$p.value


# hypothesis testing using infer package
# Calculate observed statistic
diff_mean_salary <- omega %>% 
  specify(salary ~ gender) %>% 
  calculate(stat = "diff in means", 
            order = c("female", "male")) 
diff_mean_salary

# Generate null distribution
gender_salary_null <- omega %>%
  specify(salary ~ gender) %>%
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "diff in means", 
            order = c("female", "male"))

# Visualize null distribution
gender_salary_null %>% 
  visualize() + 
  shade_p_value(obs_stat = diff_mean_salary, direction = "both") +
  labs(x = "Difference between salaries' means of female and male executives", y = "Executives number", title = "Null Hypothesis and p-value")

# Calculate p-value of observed difference
gender_salary_null %>%
  get_p_value(obs_stat = diff_mean_salary, direction = "both")

```

> What can you conclude from your analysis? A couple of sentences would be enough

> We can conclude that, at a 95% condidence level, the mean of male's salary is different from female's. There is a 95% chance that male executives earn between $4420 and $12973 higher than their female counterparts. There is a 0.0165% chance to see a difference of at least as large as $8696 between male's and female's pays.
So, we can say that the null-hypothesis is untrue, considering we have a p-value below 0.5 and a red line considerably far from 0.

## Relationship Experience - Gender?

At the board meeting, someone raised the issue that there was indeed a substantial difference between male and female salaries, but that this was attributable to other reasons such as differences in experience. A questionnaire send out to the 50 executives in the sample reveals that the average experience of the men is approximately 21 years, whereas the women only have about 7 years experience on average (see table below).

```{r, experience_stats}
# Summary Statistics of salary by gender
favstats (experience ~ gender, data = omega)

gender_exp <- omega %>%
  group_by(gender) %>%
  summarise(Mean = mean(experience), 
            SD = sd(experience), 
            Count = n(), 
            T_critical = qt(p = 0.975, df = Count-1), 
            SE = SD/sqrt(Count),
            Margin = T_critical * SE,
            Lower = Mean - Margin,
            Upper = Mean + Margin)
gender_exp
```

```{r, hypothesis test on experience differences}

# hypothesis testing using t.test() 

t_gender <- t.test(experience ~ gender, data = omega)
t_gender
t_gender$p.value

# hypothesis testing using infer package
# Calculate observed statistic
diff_mean_exp <- omega %>% 
  specify(experience ~ gender) %>% 
  calculate(stat = "diff in means", 
            order = c("female", "male")) 
diff_mean_exp

# Generate null distribution
gender_exp_null <- omega %>%
  specify(experience ~ gender) %>%
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "diff in means", 
            order = c("female", "male"))

# Visualize null distribution
gender_exp_null %>% 
  visualize() + 
  shade_pvalue(obs_stat = diff_mean_exp, direction = "both")

# Calculate p-value of observed difference
gender_exp_null %>%
  get_pvalue(obs_stat = diff_mean_exp, direction = "both")

```

Based on this evidence, can you conclude that there is a significant difference between the experience of the male and female executives? Perform similar analyses as in the previous section. Does your conclusion validate or endanger your conclusion about the difference in male and female salaries?  

> We can conclude that, at a 95% condidence level, the mean of male's experience is different from female's. There is a 95% chance that male executives have between 19.35 years and 8.13 years more experience than their female counterparts. There is a 1.23e-05 chance to see a difference of at least as large as 13.7 years between male's and female's experiences.
So, we can say that the Null-hypothesis is untrue, given we have a p-value below o.5 and a red line considerably far from 0.

## Relationship Salary - Experience ?

Someone at the meeting argues that clearly, a more thorough analysis of the relationship between salary and experience is required before any conclusion can be drawn about whether there is any gender-based salary discrimination in the company.

Analyse the relationship between salary and experience. Draw a scatterplot to visually inspect the data

```{r, salary_exp_scatter, fig.width=10}

ggplot(omega, mapping = aes(x = experience, y = salary)) + 
  geom_point() + 
  geom_smooth() +
  theme_bw() +
  facet_wrap(~gender) +
  labs(x = "Years of experience", y = "Salary", title = "Relationship between salary and experience")
```

## Check correlations between the data
You can use `GGally:ggpairs()` to create a scatterplot and correlation matrix. Essentially, we change the order our variables will appear in and have the dependent variable (Y), salary, as last in our list. We then pipe the dataframe to `ggpairs()` with `aes` arguments to colour by `gender` and make ths plots somewhat transparent (`alpha  = 0.3`).

```{r, ggpairs, fig.width=10}
omega %>% 
  select(gender, experience, salary) %>% #order variables they will appear in ggpairs()
  ggpairs(aes(colour=gender, alpha = 0.3))+
  theme_bw()
```

> Look at the salary vs experience scatterplot. What can you infer from this plot? Explain in a couple of sentences

> It looks like salary is positively correlated with experience for both genders. But as the correlation figures suggest, the positive correlation is stronger for female than for male employees.
Finally, given that correlation between salary and experience is 0.803, the salary gap between the two groups may be due to the experience difference among genders instead of being a result of gender discrimination.